---
title: "Pivot Lobsters"
author: "Matthew Pfaff"
date: "1/27/2020"
output: html_document
---

Attaching our packages and reading in data

```{r}
library(tidyverse)
library(readxl)
library(here)
library(skimr) # install.packages("skimr") 
library(kableExtra) # install.packages("kableExtra")

# read in lobster data

lobsters <- read_xlsx(here("data/lobsters2.xlsx"),
                      skip = 4)

```


Viewing our data with skimr

```{r}
# these are equivalent
skimr::skim(lobsters) #explicitly stated from skimr
# skim(lobsters) # not explicity, but okay because we've loaded the package
```

# Pivot Tables

In R, pivot tables are: `group_by() %>% summarize()`
`n()` function will count observations

```{r}

# Group by year

lobsters %>% 
  group_by(year) %>% 
  summarize(count_by_year = n())

# Group by year and site

lobsters %>% 
  group_by(year, site) %>% 
  summarize(count_by_year_site = n())

# Group by site and year

lobsters %>% 
  group_by(site, year) %>% 
  summarize(count_by_site_year = n())

```

Summarize by multiple variables

```{r}

# `na.rm` is a way to remove missing values; it allows the mean to be calculated even when there is missing values

siteyear_summary <- lobsters %>% 
  group_by(site, year) %>% 
  summarize(count_by_site_year = n(),
            mean_size_mm = mean(size_mm, na.rm = TRUE),
            sd_size_mm = sd(size_mm, na.rm = TRUE))
```

Table formatting with `kableExtra::kable()`

```{r}
siteyear_summary %>%
  kable()
```

If you add in r in front of a field, it will be included in the markdown report: 

There are `r nrow(lobsters)` total lobsters including in this report.


## Activity
1. Build from our analyses: calculate the median lobster size for each site and year (hint: `size_mm and a function called `median()`)

2. Make a simple ggplot and save.

3. sync this to github

```{r}
lobsters %>% 
  group_by(site, year) %>% 
  summarize(count_by_site_year = n(),
            median_size_mm = median(size_mm, na.rm = TRUE))
```

