---
title: "Creating graphs with ggplot2"
author: "Matthew Pfaff"
date: "1/27/2020"
output: html_document
---

# Plots with ggplot2

We are going to make plots in R using the ggplot2 package from the tidyverse.

We will install these packages:

- tidyverse: suite of packages for wrangling and visualization (ggplot)
- readxl: to read in excel files (in many ways, shapes and sizes)
- here: to help with filepaths

```{r setup, message=FALSE}
library(tidyverse) # install.packages("tidyverse")
library(readxl) # install.packages("readxl")
library(here) # install.packages("here")
```

# California National Parks Data

```{r read data}

ca_np <- read_csv(here("data/ca_np.csv"))

ci_np <- read_excel(here("data/ci_np.xlsx"))

```


### Make my own first graph

We're going to make a line plot of year (x-variable) v. visitors (y-variable) in ggplot:

Insert code chunk = CTRL + ALT + I

```{r}

ggplot(data = ci_np, aes(x= year, y = visitors)) + 
  geom_area()

# Update the geom_ to geom_point(), geom_col(), geom_area()

```

Making a base layer for channel islands plots

```{r}
ggbase <- ggplot(data = ci_np, aes(x= year, y = visitors))
```

### Now let's start customizing!

- Line/point colors: color
- Change line type: linetype
- Change point style: pch
- Change transparency: alpha

```{r}
ggbase +
  geom_line(
    color = "purple",
    linetype = "dashed"
  )
```

Create a ggplot scatterplot (geom_point) that has triangle symbols(hint: ?pch), of a non-default size (size = #), and that are an updated color

```{r}
ggbase +
  geom_point(
    pch = 2,
    size = 4,
    color = "dodgerblue"
  )
```

### Update graph aesthetics based on variables and try adding an argument to update the color based on the year value

```{r}
ggbase +
  geom_point(
    aes(size = visitors, color = year),
    pch = 10
  )
```

