---
title: 'R Workshop Day 2: Tidying Session'
author: "Matthew Pfaff"
date: "1/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)

# Attaching packages

library(tidyverse) # install.packages("tidyverse")
library(janitor) # install.packages("janitor")
library(here) # install.packages("here")
library(readxl) # install.packages("readxl")
```

### Read in inverts.xlsx

```{r}
inverts <- read_excel(here("data/inverts.xlsx"))
```

### Pivot Longer

`pivot_longer` is how we will convert wide date to long data. We can use a single colon (:) to indicuate a series of sequential columns (e.g., A:D).

What if you want some in series and some NOT in series?
c('2016':'2018', '2020')

```{r}
inverts_long <- pivot_longer(data = inverts, 
                             cols = '2016':'2018',
                             names_to = "year",
                             values_to = "species_count")
```

### Pivot Wider

`pivot_wider` is how we can convert long data to wide data. We are using the pipe operator here ( %>% AKA "and then") to show another way of doing it.

```{r}
inverts_wide <- inverts_long %>% 
  pivot_wider(names_from = common_name,
              values_from = species_count)
```

### Cleaning Names

`janitor::clean_names()`

We don't like spaces in data because you have to have quotes around the field. We can use the Janitor package to clean the names into whatever we want.

```{r}
inverts_wide <- inverts_wide %>%
  clean_names()
```

### Combining information from columns

`tidyr::unite()` to combine information from columns

Example: merge the site code and year into a signle column called "site_year"

```{r}
inverts_unite <- inverts_long %>% 
  unite(col = "site_year",
        c(site, year),
        sep = "_")
```

### Activity

Create a new object called `inverts_moyr` starting from inverts_long that unites the month and year columns into a single column name "mo_yr", using a slash "/" as the separator ... then try some other fun separator.

```{r}
inverts_moyr <- inverts_long %>%
  unite(col = "mo_yr",
        c(month, year),
        sep = "/")
```

### Separating information from a column

`tidyr::separate()` to separate information from a column

```{r}
inverts_sep <- inverts_unite %>% 
  separate(col = site_year,
           into = c("my_site", "my_year"))
```

### How do we work with words in our data

`stringr`

`mutate()` allows you to add a column while keeping existing columns or overwrite an existing column (but be careful with overwriting because it is generally better to add a column rather than delete a column)

Example: using built in `trees` dataset with mutate() to add 5 to all the tree heights

```{r}
trees_new <- trees %>% 
  mutate(height_update = Height + 5)
```








